---
title: "Using GeoDeepDive (xDD) to Ping the White and Grey Literature for Parrot Management"
date: "Last updated: `r Sys.Date()`"
output: github_document
---
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-lightgrey.svg)](https://www.tidyverse.org/lifecycle/#experimental)

## About
This project seeks to identify publications within which keywords related to parrot management or parrot culling has been referenced or mentioned. This project uses pre-extracted information available via API from the [GeoDeepDive](https://geodeepdive.org) infrastructure. 

Note: This readme is difficult, I know, but is not currenlty intended for a broad audience. 
Note2: Reproduction requires some knowledge or patience with Jupyter and R. 


```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
## if you open the .rproj you should be in the right directory (repository root)

# source the R function(s)
source("R/get_xdd_df.R")

# load packages
  library(ggpubr)
  library(tidyverse)
  library(jsonlite)
  library(rjson)

# plotting themes and settings
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
theme_set(ggpubr::theme_classic2())
```

## Hit up the GeoDeepDive (xDD) API using a Jupyter Notebook

Open the Jupyter notebook in your web browser by doing the following:  
1. Open terminal (Mac) or Command Line (Windows, Linux)  
2. Navigate to this directory (`cd ~/path/to/parrot_xdd/`)  
3. Open the .ipynb (type: `jupyter notebook`).  This opens the notebook directory in your web browser.   
4. Navigate to (i.e. click) the file `parrot_xDD.ipynb`: this will open the notebook and a Kernel.  
5. Edit the terms (keywords, search phrases). Edit the code chunk containing terms. See section: `Set a term or loop through terms`, and edit inside brackets at `terms = [...]`  
6. Save the notebook (it typically autosaves but just to be safe). Ctrl+S (MS) or Cmd+S (Mac) or >File>Save and Checkpoint   
7. Execute the notebook (Run or >Kernel>Restart and Run All)  
8. Navigate to the results (~/xdd_json/)  


## Import the JSON results to R
```{r filestoimport, message=FALSE, warning=FALSE,echo=TRUE}
# Import the JSON files and munge into a single data frame ----------------
files <- list.files(here::here("xdd_json"), pattern='.txt', full.names=TRUE)
# Identify and delete the empty .txt files. Yes, this is poor practice but oh well!
for(i in seq_along(files)){
  if(i==1) keep <-NULL
if(file.info(files[i])$size > 5){ # retrieve file only if it's >5bytes.
  keep <- c(keep,i)}
  }
```

Note that the following files, associated with search terms, did not yield any results:
`r files[keep]`


## Import and munge the xdd json into an R df
```{r import, echo=TRUE, message=FALSE, warning=FALSE}
parrot_xdd <- get_xdd_df(files[keep])
parrot_unique <- parrot_xdd %>% distinct(doi, .keep_all = TRUE) %>% dplyr::select(-searchterm)
```

## Overview of parrot records (`r paste("N= ", nrow(parrot_unique))`) retrieved from the GeoDeepDive database.
We have used various search terms to identify potential works of relevance to parrot management. Here we found only __`r parrot_xdd %>% dplyr::select(-filename, -searchterm, -highlight) %>% n_distinct()`__ unique publications which used our terms (see ~/parrot_xDD.ipynb). 

<!-- Exact search phrase | Number of unique records in xDD/GDD  -->
<!-- --------------|---- -->
<!-- "management of parrot*" | `r parrot_xdd %>% filter(grepl("management of parrot", searchterm, ignore.case = TRUE)) %>% dplyr::select(-filename, -highlight) %>% n_distinct()` -->
<!-- "parrot control" | `r parrot_xdd %>% filter(grepl("parrot control", searchterm, ignore.case = TRUE)) %>% dplyr::select(-filename, -highlight) %>% n_distinct()` -->
<!-- "parrot management" | `r parrot_xdd %>% filter(grepl("parrot management", searchterm, ignore.case = TRUE)) %>% dplyr::select(-filename, -highlight) %>% n_distinct()` -->
```{r papertable, echo=FALSE}
kableExtra::kable(parrot_unique %>% dplyr::select(title, coveryear, pubname, doi, URL))
```

```{r pubsvtime, echo=FALSE, eval=FALSE}
ggplot(parrot_unique %>% group_by(searchterm, coveryear) %>% 
         summarise(total_pubs = n()) %>% na.omit()) +
  geom_point(aes(x=coveryear, y=total_pubs, color=searchterm))+
  scale_color_manual(values=cbPalette)+
  labs(x="year", y="total records")+
  theme(legend.title=element_blank(), legend.position = "top")
``` 

<!-- Distribution of  records among unique publishers. -->
```{r topjrnls, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
tmp <-  parrot_xdd %>% dplyr::select(-filename, -searchterm) %>% distinct()
tmp <- bind_rows(tmp, parrot_xdd %>% dplyr::select(-filename, -searchterm)%>% filter(is.na(doi)) %>% distinct())

ggplot(tmp %>% group_by(pubname) %>% 
         summarise(n = n())) +
  geom_histogram(aes(n))+
  labs(x="number of publishers", y="total records")
```

Distribution across publications:
```{r topjrnls2, echo=FALSE, message=FALSE, warning=FALSE}
ggdotchart(
  parrot_unique %>% group_by(pubname) %>%
    summarise(n = n()) ,
  x = "pubname",
  y = "n",
  rotate = TRUE,
  ylab = "number of records",
  xlab = "",
  sorting = "desc",
  ggtheme = theme_minimal()
)
```

## Results Files
Overviews of the results can be found in the [results](/results) folder in this repository.
```{r write_outputs, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
dir.create("out")
write.csv(parrot_unique, "out/unique_results.csv")
```

### Total records per publication
Here is a comprehensive list of all publications with relevant terms (except "BBL"):
```{r table_pubs, echo=FALSE, warning=FALSE, message=FALSE}
library(kableExtra)
parrot_xdd %>% distinct(coveryear, pubname, publisher, title, authors, searchterm)%>%
  group_by(pubname) %>% 
  summarise(count = n()) %>% 
  na.omit(pubname) %>% 
  kable(caption = "Number of records (count) per publication name (pubname)", format="markdown") %>%
  kable_styling()
````

## More Information
### This project relies on the UW Deepdive Infrastructure
[See the UW deepdive infrastructure](UW-Deepdive-Infrastructure/app-template) for more details. Check out [the
wiki](https://github.com/UW-Deepdive-Infrastructure/app-template/wiki)
for more information on getting started.

### Disclaimer
This software is preliminary or provisional and is subject to revision. It is being provided to meet the need for timely best science. The software has not received final approval by the U.S. Geological Survey (USGS). No warranty, expressed or implied, is made by the USGS or the U.S. Government as to the functionality of the software and related material nor shall the fact of release constitute any such warranty. The software is provided on the condition that neither the USGS nor the U.S. Government shall be held liable for any damages resulting from the authorized or unauthorized use of the software.
